---
title: "Per Second Heart Rate Data Sampling and Merging"
output: html_notebook
---

Created a 20% sample of each heart rate csv file via: 

`Rscript save_sample_csv.R 0.2 20170801_20180315_heartrate_seconds_merged.csv 20180316_20180701_heartrate_seconds_merged.csv 20180702_20180927_heartrate_seconds_merged.csv`

(`save_sample_csv.R` is included in this directory)

Merged all sample heart rate files via the command line:

`cat 20170801_20180315_heartrate_seconds_merged_sample.csv 20180316_20180701_heartrate_seconds_merged_sample.csv 20180702_20180927_heartrate_seconds_merged_sample.csv > 20170801_20180927_heartrate_seconds_merged_sample_mergedAll.csv`

```{r}
library(data.table)
library(magrittr)
```

## Per Second Heart Rate Data Overview

```{r}
dt_hr <- fread("data/20170801_20180927_heartrate_seconds_merged_sample_mergedAll.csv")
```

```{r}
dt_hr %>% head()
```

Number of participants (number of unique Ids):
```{r}
dt_hr$Id %>% unique() %>% length()
```

There are no missing heart rate values:
```{r}
dt_hr$Value %>% is.na() %>% sum() %>% divide_by(nrow(dt_hr))
```




