---
title: "Per Second Heart Rate Data Sampling"
output: html_notebook
---

The per second heart rate data starts on 08/01/2017 and ends on 09/27/2018 (423 days, inclusive of start and end dates). It's divided into 3 files:

* `20170801_20180315_heartrate_seconds_merged_sample.csv`
* `20180316_20180701_heartrate_seconds_merged_sample.csv` 
* `20180702_20180927_heartrate_seconds_merged_sample.csv`

This sampling will retrieve the first 85-day (20% of 423 is ~85) contiguous block of measurements for each participant. 

Different participants have measurements starting on different dates and lasting for different amounts of time. Sampling the first 85-day block of measurements relative to each participant's start date will create some consistency. Participants that have significantly less measurements (e.g. they don't have measurements for the whole 85-day block) will later be removed.

This sampling will:

1. Concatenate all 3 files into one large data.table structure.
2. Remove measurements that occur during night times (10pm to 6am).
3. Sample amd save only the first 85-day contiguous block of measurements for each participant.

See `sample.R` for code.

```{r}
library(data.table)
library(magrittr)
```

## Sampled Per Second Heart Rate Data Overview

```{r}
load("RData_intermediate/allHR_85days.RData")
```


```{r}
dt_sample
```

Number of participants (number of unique Ids):
```{r}
dt_sample$Id %>% unique() %>% length()
```

Number of measurements per participant:
```{r}
dt_sample[, .N, by=Id]
# note: participants with too few measurements will be removed later
```

There are no missing heart rate values:
```{r}
dt_sample$Value %>% is.na() %>% sum() %>% divide_by(nrow(dt_sample))
```




